---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  warning = FALSE, 
  message = FALSE, 
  collapse = TRUE,
  comment = "#>"
)
setwd("D:/CU/Thesis/BinaryShapeClust/BinaryShapeClust")
```

# Binary Shape-Based Clustering

<!-- badges: start -->
<!-- badges: end -->

This repository is accompanies "Shape-Based Clustering of Daily Weigh-In Trajectories using Dynamic 
Time Warping" (doi). The aim of the paper was to use Dynamic Time Warping, a shape-based clutering method, to cluster binary trajectories and evaluate patterns. Study data is not made publically available, however simulated data to resemble the study data is used and shared in this repository. 

- Source files to load in data (data cleaning and simulation data)       

```{r}
source("Code/01_CleanStudyData.R")
source("Code/02_SimulateClusters.R")
```

- Plot the binary heatmap (this should be a file of its own to make heat maps based on the data) 

```{r, fig.width = 15}
#### Function not incorporated yet
source("Code/ClusterHeatmap.R")

ggplot(sim_plot, aes(x = day, y = participant_id, fill = weighed_in)) +
    geom_tile() + ggtitle("True Data Clusters and Mean Adherence %") + 
    theme(plot.title = element_text(hjust = 0.5)) + 
    scale_fill_viridis(discrete = TRUE, option="A") +
    facet_wrap(~cluster_assignment, scales = "free", ncol = 5)
```


- Cluster the data using Euclidean, Jaccard, and DTW and show how to use probability windows and linkages
